### 2016-04-03 Proposal/refs.bib.tex

@incollection{NIPS2015_5945,
author = {Hermann, Karl Moritz and Kocisky, Tomas and Grefenstette, Edward and Espeholt, Lasse and Kay, Will and Suleyman, Mustafa and Blunsom, Phil},
booktitle = {Advances in Neural Information Processing Systems 28},
editor = {Cortes, C and Lawrence, N D and Lee, D D and Sugiyama, M and Garnett, R},
pages = {1684--1692},
publisher = {Curran Associates, Inc.},
title = {{Teaching Machines to Read and Comprehend}},
url = {http://papers.nips.cc/paper/5945-teaching-machines-to-read-and-comprehend.pdf},
year = {2015}
}